---
title: "Incubation vs chick-rearing"
author: "Gemma Clucas"
date: "4/19/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(tidyverse)
library(viridis)
library(knitr)
```

### Read in the rarefied feature table
This has had taxonomy corrections made e.g. grouping all River herring.
```{r}
df <- read.csv("MiFish/WSC3_terns/Data-for-WSC3-only_plates7-14.csv", header = TRUE) 
```

Read in table with full sample info. I will use this to pull out the chick-rearing/incubation information.
```{r}
meta_df <- read.csv("MiFish/WSC3_terns/2019_COTE_all_metadata.csv", header = TRUE) %>% 
  rename(BreedingPeriod = Breeding.period..incubation..Chick.rearing..Mixed.)
```

Add the incubation vs chick-rearing information to the feature table.
```{r}
df <- meta_df %>% 
  select(Unique_ID, BreedingPeriod, Date) %>% 
  rename(index = Unique_ID) %>% 
  left_join(df, ., by = "index") 
```

### Some functions to calculate frequency of occurrence
```{r}
# A function to calculate FOO given a filtered feature table and the number of samples
calc_FOO <- function(x, n_samples) {
  x %>% 
  mutate_if(is.numeric, ~1 * (. > 0)) %>%     # change to detection/non-detection
  summarise_each(funs = sum) %>%              # count number of detections
  melt() %>%                                  # make into long dataframe
  rename(Occurrence = value,
         Species = variable) %>% 
  mutate(FOO = Occurrence/n_samples*100)
}


# get rid of periods in column names and order species
scrub_periods <- function(x) {
  x$Species <-  gsub("\\.", " ", x$Species)
  # x$Species <- factor(x$Species, levels = Species_ordered)
}

# combine the first two functions but also allow to discriminate chick-rearing or incubation
do_everything <- function(Colony_filter, Age_filter){
  x <- df %>% 
    filter(Colony == Colony_filter & Age == Age_filter) %>% 
    select(River.herring:Rainbow.smelt)
  n_samples <- nrow(x)
  y <- calc_FOO(x, n_samples)
  y$Species <- scrub_periods(y)
  return(y)
}


# combine the first two functions but also allow to discriminate chick-rearing or incubation
do_everything2 <- function(Colony_filter, Age_filter, BreedingPeriod_filter){
  x <- df %>% 
    filter(Colony == Colony_filter & Age == Age_filter & BreedingPeriod == BreedingPeriod_filter) %>% 
    select(River.herring:Rainbow.smelt)
  n_samples <- nrow(x)
  y <- calc_FOO(x, n_samples)
  y$Species <- scrub_periods(y)
  return(y)
}

```

### Some FOO calculations
I know there must be a way to use map() to do this neatly, but my brain cannot figure it out right now.
```{r}
FOO_MON_CHICK <- do_everything("Monomoy NWR", "CHICK") %>% mutate(Colony = "Monomoy NWR", Age = "CHICK")
FOO_MON_ADULT <- do_everything("Monomoy NWR", "ADULT") %>% mutate(Colony = "Monomoy NWR", Age = "ADULT")
FOO_SEA_CHICK <- do_everything("White and Seavey Is.", "CHICK") %>% mutate(Colony = "White and Seavey Is.", Age = "CHICK")
FOO_SEA_ADULT <- do_everything("White and Seavey Is.", "ADULT") %>% mutate(Colony = "White and Seavey Is.", Age = "ADULT")
FOO_STI_CHICK <- do_everything("Stratton Is.", "CHICK") %>% mutate(Colony = "Stratton Is.", Age = "CHICK")
FOO_STI_ADULT <- do_everything("Stratton Is.", "ADULT") %>% mutate(Colony = "Stratton Is.", Age = "ADULT")
FOO_OGI_CHICK <- do_everything("Outer Green Is.", "CHICK") %>% mutate(Colony = "Outer Green Is.", Age = "CHICK")
FOO_OGI_ADULT <- do_everything("Outer Green Is.", "ADULT") %>% mutate(Colony = "Outer Green Is.", Age = "ADULT")
FOO_JEN_CHICK <- do_everything("Jenny Is.", "CHICK") %>% mutate(Colony = "Jenny Is.", Age = "CHICK")
FOO_JEN_ADULT <- do_everything("Jenny Is.", "ADULT") %>% mutate(Colony = "Jenny Is.", Age = "ADULT")
FOO_PON_CHICK <- do_everything("Pond Is.", "CHICK") %>% mutate(Colony = "Pond Is.", Age = "CHICK")
FOO_PON_ADULT <- do_everything("Pond Is.", "ADULT") %>% mutate(Colony = "Pond Is.", Age = "ADULT")
FOO_EGG_CHICK <- do_everything("Eastern Egg Rock", "CHICK") %>% mutate(Colony = "Eastern Egg Rock", Age = "CHICK")
FOO_EGG_ADULT <- do_everything("Eastern Egg Rock", "ADULT") %>% mutate(Colony = "Eastern Egg Rock", Age = "ADULT")
FOO_MET_CHICK <- do_everything("Metinic", "CHICK") %>% mutate(Colony = "Metinic", Age = "CHICK")
FOO_MET_ADULT <- do_everything("Metinic", "ADULT") %>% mutate(Colony = "Metinic", Age = "ADULT")
FOO_SINWR_CHICK <- do_everything("Seal Is. NWR", "CHICK") %>% mutate(Colony = "Seal Is. NWR", Age = "CHICK")
FOO_SINWR_ADULT <- do_everything("Seal Is. NWR", "ADULT") %>% mutate(Colony = "Seal Is. NWR", Age = "ADULT")
FOO_SHI_CHICK <- do_everything("Ship Is.", "CHICK") %>% mutate(Colony = "Ship Is.", Age = "CHICK")
FOO_SHI_ADULT <- do_everything("Ship Is.", "ADULT") %>% mutate(Colony = "Ship Is.", Age = "ADULT")
FOO_PMI_CHICK <- do_everything("Petit Manan Is.", "CHICK") %>% mutate(Colony = "Petit Manan Is.", Age = "CHICK")
FOO_PMI_ADULT <- do_everything("Petit Manan Is.", "ADULT") %>% mutate(Colony = "Petit Manan Is.", Age = "ADULT")
```

Do the same but split adult samples into incubation and chick-rearing.
```{r}

do_everything2("Monomoy NWR", "ADULT", "Mixed") %>% 
  mutate(Colony = "Monomoy NWR", Age = "ADULT", Period = "Mixed") %>% 
  bind_rows(., do_everything2("White and Seavey Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "White and Seavey Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("White and Seavey Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "White and Seavey Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Stratton Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Stratton Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Stratton Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Stratton Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Outer Green Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Outer Green Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Outer Green Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Outer Green Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Jenny Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Jenny Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Jenny Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Jenny Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Pond Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Pond Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Pond Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Pond Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Eastern Egg Rock", "ADULT", "incubation") %>% 
              mutate(Colony = "Eastern Egg Rock", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Eastern Egg Rock", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Eastern Egg Rock", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Metinic", "ADULT", "incubation") %>% 
              mutate(Colony = "Metinic", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Metinic", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Metinic", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Seal Is. NWR", "ADULT", "incubation") %>% 
              mutate(Colony = "Seal Is. NWR", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Seal Is. NWR", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Seal Is. NWR", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Ship Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Ship Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Ship Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Ship Is.", Age = "ADULT", Period = "Chick-rearing")) %>% 
  bind_rows(do_everything2("Petit Manan Is.", "ADULT", "incubation") %>% 
              mutate(Colony = "Petit Manan Is.", Age = "ADULT", Period = "Incubation")) %>% 
  bind_rows(do_everything2("Petit Manan Is.", "ADULT", "chick-rearing") %>% 
              mutate(Colony = "Petit Manan Is.", Age = "ADULT", Period = "Chick-rearing"))
```
### Make some plots


