---
title: "Plots for WSC3"
author: "Gemma Clucas"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(tidyverse)
library(viridis)
library(knitr)
```

### Read in the rarefied feature table

```{r}
df <- read.csv("MiFish/WSC3_terns/Data-for-WSC3-only_plates7-14.csv", header = TRUE) 
```

### Create plotting order
I need to define the order of the fish along the x-axis, so that it is consistent among graphs.

```{r}
# Define x axis order
Species_ordered <-  c("Atlantic herring",
                      "River herring",
                      "Sandlance sp",
                      "White hake",
                      "Silver hake",
                      "Red hake",
                      "Pollock",
                      "Fourbeard rockling",
                      "Haddock",
                      "Atlantic cod",
                      "Atlantic mackerel",
                      "Atlantic butterfish",
                      "Cunner",
                      "Redfish sp",
                      "Mummichog",
                      "Atlantic silverside",
                      "Atlantic tomcod",
                      "Spotted codling",
                      "Three spined stickleback",
                      "Black spotted stickleback",
                      "Nine spine stickleback",
                      "Radiated shanny",
                      "American angler",
                      "Tautog",
                      "Darter sp",
                      "Rock gunnel",
                      "Sculpin sp",
                      "Cusk",
                      "Atlantic menhaden",
                      "Lumpfish",
                      "Northern searobin",
                      "Carangidae indet",
                      "Red lionfish",
                      "Rainbow smelt",
                      "Chub mackerel")

# # These colours are the ones Liz originally had, although I have messed with the order somewhat
# Species_with_colours <- c("Atlantic herring" = rgb(98, 64, 18, max = 255),
#                       "River herring" = rgb(150, 102, 34, max = 255),
#                       "Sandlance sp" = rgb(207, 138, 54, max = 255),
#                       "White hake" = rgb(247, 224, 167, max = 255),
#                       "Silver hake" = rgb(225, 204, 150, max = 255),
#                       "Red hake" = rgb(187, 169, 126, max = 255),
#                       "Pollock" = rgb(214,	237, 234, max = 255),
#                       "Fourbeard rockling" = rgb(159, 211, 204, max = 255),
#                       "Haddock" = rgb(92,	164, 160, max = 255),
#                       "Atlantic cod" = rgb(50,	118, 113, max = 255),
#                       "Atlantic mackerel" = rgb(30, 75, 63, max = 255),
#                       "Atlantic butterfish" = rgb(5, 60, 245, max = 255),
#                       "Cunner" = rgb(5, 57, 237, max = 255),
#                       "Redfish sp" = rgb(6, 56, 234, max = 255),
#                       "Mummichog" = rgb(4, 50, 211, max = 255),
#                       "Atlantic silverside" = rgb(3, 44, 190, max = 255),
#                       "Atlantic tomcod" = rgb(3, 46, 199, max = 255),
#                       "Spotted codling" = rgb(3, 39, 175, max = 255),
#                       "Three spined stickleback" = rgb(1,	34,	157, max = 255),
#                       "Black spotted stickleback" = rgb(2, 29, 141, max = 255),
#                       "Nine spine stickleback" = rgb(1, 26, 122, max = 255),
#                       "Radiated shanny" = rgb(2, 21, 100, max = 255),
#                       "American angler" = rgb(2,	19,	90, max = 255),
#                       "Tautog" = rgb(1,	15,	80, max = 255),
#                       "Darter sp" = rgb(0, 8, 60, max = 255),
#                       "Rock gunnel" = rgb(0, 0,0, max = 255),
#                       "Sculpin sp"= rgb(0, 0,0, max = 255),
#                       "Cusk"= rgb(0, 0,0, max = 255),
#                       "Atlantic menhaden"= rgb(0, 0,0, max = 255),
#                       "Lumpfish"= rgb(0, 0,0, max = 255),
#                       "Northern searobin"= rgb(0, 0,0, max = 255),
#                       "Carangidae indet." = rgb(0, 0,0, max = 255),
#                       "Rainbow smelt" = rgb(0, 0,0, max = 255),
#                       "Chub mackerel" = rgb(0, 0,0, max = 255))

# Here I am going to make everything after haddock grey.

grey <- rgb(200,	200, 200, max = 255)

Species_with_colours <- c("Atlantic herring" = rgb(98, 64, 18, max = 255),
                      "River herring" = rgb(150, 102, 34, max = 255),
                      "Sandlance sp" = rgb(207, 138, 54, max = 255),
                      "White hake" = rgb(247, 224, 167, max = 255),
                      "Silver hake" = rgb(225, 204, 150, max = 255),
                      "Red hake" = rgb(187, 169, 126, max = 255),
                      "Pollock" = rgb(214,	237, 234, max = 255),
                      "Fourbeard rockling" = rgb(159, 211, 204, max = 255),
                      "Haddock" = rgb(92,	164, 160, max = 255),
                      "Atlantic cod" = grey,
                      "Atlantic mackerel" = grey,
                      "Atlantic butterfish" = grey,
                      "Cunner" = grey,
                      "Redfish sp" = grey,
                      "Mummichog" = grey,
                      "Atlantic silverside" = grey,
                      "Atlantic tomcod" = grey,
                      "Spotted codling" = grey,
                      "Three spined stickleback" = grey,
                      "Black spotted stickleback" = grey,
                      "Nine spine stickleback" = grey,
                      "Radiated shanny" = grey,
                      "American angler" = grey,
                      "Tautog" = grey,
                      "Darter sp" = grey,
                      "Rock gunnel" = grey,
                      "Sculpin sp"= grey,
                      "Cusk"= grey,
                      "Atlantic menhaden"= grey,
                      "Lumpfish"= grey,
                      "Northern searobin"= grey,
                      "Carangidae indet." = grey,
                      "Rainbow smelt" = grey,
                      "Chub mackerel" = grey)


```

### Functions to calculate FOO from filtered feature table and plot FOO
I am going to filter the feature table for e.g. COTE adults from 2017 by hand, but then I just want to be able to use functions to calculate FOO and plot, to save on the amount of code I have to write.

```{r}

calc_FOO <- function(x, n_samples) {
  x %>% 
  mutate_if(is.numeric, ~1 * (. > 0)) %>%     # change to detection/non-detection
  summarise_each(funs = sum) %>%              # count number of detections
  melt() %>%                                  # make into long dataframe
  rename(Occurrence = value,
         Species = variable) %>% 
  mutate(FOO = Occurrence/n_samples*100)
}


# get rid of periods in column names and order species
scrub_periods <- function(x) {
  x$Species <-  gsub("\\.", " ", x$Species)
  x$Species <- factor(x$Species, levels = Species_ordered)
}

cols <- c(rgb(142, 202, 230, max = 255), "#8ECAE6", "#219EBC", "#023047", "#FFB703", "#FB8500")

# function to plot, with the species in the order determined by Species_with_colours
plot_FOO <- function(x) {
  x %>% 
    ggplot() +
    geom_bar(aes(x = Species, y = FOO), stat = "identity", fill = cols[1]) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45,  hjust=1)) +
    labs(y = "Frequency of occurrence (%)") +
    ylim(c(0, 100))
    #scale_fill_manual(values = cols[1])
}

```

### Monomoy chicks
Select the Monomoy chick data using ```filter```, change to presence/absence data using the ```mutate_if``` term, then ```melt``` to calculate the frequence of occurrence.

```{r, message=FALSE}
# Filter the data and save to object
MON_chicks <- df %>% 
  filter(Colony == "Monomoy NWR" & Age == "CHICK") %>% 
  select(River.herring:Rainbow.smelt) 

# record number of records for calculating FOO
n_samples <- nrow(MON_chicks)
FOO <- calc_FOO(MON_chicks, n_samples)

FOO$Species <- scrub_periods(FOO)

FOO %>% 
  filter(FOO > 0) %>% 
  plot_FOO()
```


### Plotting them all together on one, multipanel figure

I think I can use faceting to do this, but I need to combine all the FOO data together into one dataframe. First, save the FOO data for each group.


```{r}

do_everything <- function(Colony_filter, Age_filter){
  x <- df %>% 
    filter(Colony == Colony_filter & Age == Age_filter) %>% 
    select(River.herring:Rainbow.smelt)
  n_samples <- nrow(x)
  y <- calc_FOO(x, n_samples)
  y$Species <- scrub_periods(y)
  return(y)
}

FOO_MON_CHICK <- do_everything("Monomoy NWR", "CHICK") %>% mutate(Colony = "Monomoy NWR", Age = "CHICK")
FOO_MON_ADULT <- do_everything("Monomoy NWR", "ADULT") %>% mutate(Colony = "Monomoy NWR", Age = "ADULT")
FOO_SEA_CHICK <- do_everything("White and Seavey Is.", "CHICK") %>% mutate(Colony = "White and Seavey Is.", Age = "CHICK")
FOO_SEA_ADULT <- do_everything("White and Seavey Is.", "ADULT") %>% mutate(Colony = "White and Seavey Is.", Age = "ADULT")
FOO_STI_CHICK <- do_everything("Stratton Is.", "CHICK") %>% mutate(Colony = "Stratton Is.", Age = "CHICK")
FOO_STI_ADULT <- do_everything("Stratton Is.", "ADULT") %>% mutate(Colony = "Stratton Is.", Age = "ADULT")
FOO_OGI_CHICK <- do_everything("Outer Green Is.", "CHICK") %>% mutate(Colony = "Outer Green Is.", Age = "CHICK")
FOO_OGI_ADULT <- do_everything("Outer Green Is.", "ADULT") %>% mutate(Colony = "Outer Green Is.", Age = "ADULT")
FOO_JEN_CHICK <- do_everything("Jenny Is.", "CHICK") %>% mutate(Colony = "Jenny Is.", Age = "CHICK")
FOO_JEN_ADULT <- do_everything("Jenny Is.", "ADULT") %>% mutate(Colony = "Jenny Is.", Age = "ADULT")
FOO_PON_CHICK <- do_everything("Pond Is.", "CHICK") %>% mutate(Colony = "Pond Is.", Age = "CHICK")
FOO_PON_ADULT <- do_everything("Pond Is.", "ADULT") %>% mutate(Colony = "Pond Is.", Age = "ADULT")
FOO_EGG_CHICK <- do_everything("Eastern Egg Rock", "CHICK") %>% mutate(Colony = "Eastern Egg Rock", Age = "CHICK")
FOO_EGG_ADULT <- do_everything("Eastern Egg Rock", "ADULT") %>% mutate(Colony = "Eastern Egg Rock", Age = "ADULT")
FOO_MET_CHICK <- do_everything("Metinic", "CHICK") %>% mutate(Colony = "Metinic", Age = "CHICK")
FOO_MET_ADULT <- do_everything("Metinic", "ADULT") %>% mutate(Colony = "Metinic", Age = "ADULT")
FOO_SINWR_CHICK <- do_everything("Seal Is. NWR", "CHICK") %>% mutate(Colony = "Seal Is. NWR", Age = "CHICK")
FOO_SINWR_ADULT <- do_everything("Seal Is. NWR", "ADULT") %>% mutate(Colony = "Seal Is. NWR", Age = "ADULT")
FOO_SHI_CHICK <- do_everything("Ship Is.", "CHICK") %>% mutate(Colony = "Ship Is.", Age = "CHICK")
FOO_SHI_ADULT <- do_everything("Ship Is.", "ADULT") %>% mutate(Colony = "Ship Is.", Age = "ADULT")
FOO_PMI_CHICK <- do_everything("Petit Manan Is.", "CHICK") %>% mutate(Colony = "Petit Manan Is.", Age = "CHICK")
FOO_PMI_ADULT <- do_everything("Petit Manan Is.", "ADULT") %>% mutate(Colony = "Petit Manan Is.", Age = "ADULT")

All_FOO <- FOO_MON_CHICK %>% 
  bind_rows(., FOO_MON_ADULT) %>% 
  bind_rows(., FOO_SEA_CHICK) %>% 
  bind_rows(., FOO_SEA_ADULT) %>% 
  bind_rows(., FOO_STI_CHICK) %>% 
  bind_rows(., FOO_STI_ADULT) %>% 
  bind_rows(., FOO_OGI_CHICK) %>% 
  bind_rows(., FOO_OGI_ADULT) %>% 
  bind_rows(., FOO_JEN_CHICK) %>% 
  bind_rows(., FOO_JEN_ADULT) %>% 
  bind_rows(., FOO_PON_CHICK) %>%
  bind_rows(., FOO_PON_ADULT) %>%
  bind_rows(., FOO_EGG_CHICK) %>%
  bind_rows(., FOO_EGG_ADULT ) %>%
  bind_rows(., FOO_MET_CHICK) %>%
  bind_rows(., FOO_MET_ADULT) %>%
  bind_rows(., FOO_SINWR_CHICK) %>%
  bind_rows(., FOO_SINWR_ADULT) %>%
  bind_rows(., FOO_SHI_CHICK) %>%
  bind_rows(., FOO_SHI_ADULT) %>%
  bind_rows(., FOO_PMI_CHICK) %>%
  bind_rows(., FOO_PMI_ADULT) 
  
All_FOO$Species <- scrub_periods(All_FOO)

## There isn't enough space to plot it as barplots with faceting, so commenting this out.
# All_FOO %>% 
#   filter(FOO > 0) %>% 
#   ggplot() +
#   geom_bar(aes(x = Species, y = FOO), stat = "identity") +
#   facet_grid(rows = vars(Age), cols = vars(Colony)) +
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 45,  hjust=1, colour = "black", margin = margin(t = 0)),
#         axis.text.y = element_text(colour = "black"),
#         axis.text=element_text(size=7),
#         axis.title=element_text(size=8, face = "bold"),
#         axis.ticks.x = element_blank(),
#         panel.border=element_rect(colour="black",size=1, fill = NA),
#         panel.spacing = unit(2, "mm"),
#         strip.background = element_blank(),
#         strip.text = element_text(colour = "black", size=8, face = "bold"),
#         axis.line = element_blank()) +
#   labs(y = "Frequency of occurrence (%)") +
#   theme(legend.position = "none")
```

Great... there is way too much to try to plot this on one figure. Can I use stacked bar charts instead?

```{r}
# Use factors for ordering the islands
All_FOO$Colony_f = factor(All_FOO$Colony, levels = c("Monomoy NWR",
                                                     "White and Seavey Is.",
                                                     "Stratton Is.",
                                                     "Outer Green Is.",
                                                     "Jenny Is.",
                                                     "Pond Is.",
                                                     "Eastern Egg Rock",
                                                     "Metinic",
                                                     "Seal Is. NWR",
                                                     "Ship Is.",
                                                     "Petit Manan Is."))

p <- All_FOO %>% 
  ggplot() +
  geom_bar(aes(x= Age, y = FOO, fill = Species), stat = "identity", position = "fill", colour = "grey40") +
  facet_grid(~Colony_f) +
  scale_fill_manual(values = Species_with_colours) +
  theme_minimal() +
  xlab("") +
  ylab("Percent occurrence")
p


ggsave(filename = "MiFish/WSC3_terns/stacked_barplots_all_colonies.jpg",
  plot = p,
  dpi = 600,
  device = "jpeg",
  width = 20,
  height = 6
)

```


What are the most frequent items across all colonies?
```{r}
All_FOO %>% group_by(Species) %>% 
  summarise(mean = mean(FOO))  %>% 
  arrange(-mean)

```


### Plotting Relative Read Abundance

```{r}
options(scipen = 999)

# To do for a single species
df %>% group_by(Colony, Age) %>% 
  summarise(reads = sum(Atlantic.herring), 
            n = n(),
            RRA = reads / (n*400))

df %>% 
  group_by(Colony, Age) %>% 
  mutate(n = n()) %>% 
  #summarise(across(c(River.herring, Rainbow.smelt)), ~ sum(.x, na.rm = TRUE))
  summarise(across(where(is.numeric) & !n, ~ sum(.x, na.rm = TRUE))) 
```

I can't find a way to sum across all the columns while keeping n unchanged or not losing. Going to have to save n in a new datafram I think, then bind.

```{r}
n_samples_per_group <- df %>% 
  group_by(Colony, Age) %>% 
  summarise(n = n()) 
```


```{r}
RRA_df <- df %>% 
  filter(!is.na(Colony)) %>% 
  group_by(Colony, Age) %>% 
  #summarise(across(c(River.herring, Rainbow.smelt)), ~ sum(.x, na.rm = TRUE))
  summarise(across(where(is.numeric) & !c(Plate, Year), ~ sum(.x, na.rm = TRUE))) %>% 
  left_join(., n_samples_per_group) %>% 
  mutate(total_reads = n*400) %>% 
  mutate(across(where(is.numeric) & !c(n,total_reads), .fns = ~./total_reads*100 ))

# melt so that it can be plotted as stacked bars
RRA_melted <- RRA_df %>% 
  select(-n, -total_reads) %>% 
  melt() %>% 
  rename(Species = variable,
         RRA = value)
  
# scrub periods and order species
RRA_melted$Species <- scrub_periods(RRA_melted)
```

Plot.Notice I'm not plotting metinic as it has small sample sizes and the graph was already large.
```{r}
# Use factors for ordering the islands
RRA_melted$Colony_f = factor(RRA_melted$Colony, levels = c("Monomoy NWR",
                                                     "White and Seavey Is.",
                                                     "Stratton Is.",
                                                     "Outer Green Is.",
                                                     "Jenny Is.",
                                                     "Pond Is.",
                                                     "Eastern Egg Rock",
                                                     "Metinic",
                                                     "Seal Is. NWR",
                                                     "Ship Is.",
                                                     "Petit Manan Is."))

r <- RRA_melted %>% 
  filter(Colony != "Metinic") %>% 
  ggplot() +
  geom_bar(aes(x= Age, y = RRA, fill = Species), stat = "identity", position = "fill", colour = "grey40") +
  facet_grid(~Colony_f) +
  scale_fill_manual(values = Species_with_colours) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.background = element_rect(fill = "transparent"), 
        legend.box.background = element_rect(fill = "transparent", colour = NA),
        strip.background =element_rect(fill="transparent")) + 
        xlab("") +
        ylab("Relative Read Abundance")
r

ggsave(filename = "MiFish/WSC3_terns/stacked_barplots_all_colonies_RRA_notMetinic.png",
  plot = r,
  dpi = 600,
  device = "png",
  width = 20,
  height = 6,
  bg = "transparent"
)


RRA_df %>% select(Colony, Age, n)
```



