---
title: "Plots for WSC3"
author: "Gemma Clucas"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(tidyverse)
library(viridis)
library(knitr)
```

### Read in the rarefied feature table

```{r}
df <- read.csv("MiFish/WSC3_terns/Data-for-WSC3-only.csv", header = TRUE) 
```

### Create plotting order
I need to define the order of the fish along the x-axis, so that it is consistent among graphs.

```{r}
# Define x axis order
Species_ordered <-  c("Atlantic herring",
                      "River herring",
                      "Sandlance",
                      "White hake",
                      "Silver hake",
                      "Red hake",
                      "Fourbeard rockling",
                      "Haddock",
                      "Saithe",
                      "Atlantic cod",
                      "Atlantic mackerel",
                      "Atlantic butterfish",
                      "Cunner",
                      "Acadian redfish",
                      "Mummichog",
                      "Atlantic silverside",
                      "Atlantic tomcod",
                      "Spotted codling",
                      "Three spined stickleback",
                      "Black spotted stickleback",
                      "Nine spine stickleback",
                      "Radiated shanny",
                      "American angler",
                      "Tautog",
                      "Darter sp",
                      "Rock gunnel",
                      "Sculpin sp",
                      "Cusk",
                      "Atlantic menhaden",
                      "Lumpfish",
                      "Northern searobin",
                      "Carangidae indet",
                      "Red lionfish")

Species_with_colours <- c("Atlantic herring" = rgb(98, 64, 18, max = 255),
                      "River herring" = rgb(150, 102, 34, max = 255),
                      "Sandlance" = rgb(207, 138, 54, max = 255),
                      "White hake" = rgb(247, 224, 167, max = 255),
                      "Silver hake" = rgb(225, 204, 150, max = 255),
                      "Red hake" = rgb(187, 169, 126, max = 255),
                      "Fourbeard rockling" = rgb(214,	237, 234, max = 255),
                      "Haddock" = rgb(159, 211, 204, max = 255),
                      "Saithe" = rgb(92,	164, 160, max = 255),
                      "Atlantic cod" = rgb(50,	118, 113, max = 255),
                      "Atlantic mackerel" = rgb(30, 75, 63, max = 255),
                      "Atlantic butterfish" = rgb(5, 60, 245, max = 255),
                      "Cunner" = rgb(5, 57, 237, max = 255),
                      "Acadian redfish" = rgb(6, 56, 234, max = 255),
                      "Mummichog" = rgb(4, 50, 211, max = 255),
                      "Atlantic silverside" = rgb(3, 44, 190, max = 255),
                      "Atlantic tomcod" = rgb(3, 46, 199, max = 255),
                      "Spotted codling" = rgb(3, 39, 175, max = 255),
                      "Three spined stickleback" = rgb(1,	34,	157, max = 255),
                      "Black spotted stickleback" = rgb(2, 29, 141, max = 255),
                      "Nine spine stickleback" = rgb(1, 26, 122, max = 255),
                      "Radiated shanny" = rgb(2, 21, 100, max = 255),
                      "American angler" = rgb(2,	19,	90, max = 255),
                      "Tautog" = rgb(1,	15,	80, max = 255),
                      "Darter sp" = rgb(0, 8, 60, max = 255),
                      "Rock gunnel" = rgb(0, 0,0, max = 255),
                      "Sculpin sp."= rgb(0, 0,0, max = 255),
                      "Cusk"= rgb(0, 0,0, max = 255),
                      "Atlantic menhaden"= rgb(0, 0,0, max = 255),
                      "Lumpfish"= rgb(0, 0,0, max = 255),
                      "Northern searobin"= rgb(0, 0,0, max = 255),
                      "Carangidae indet." = rgb(0, 0,0, max = 255))


```

### Functions to calculate FOO from filtered feature table and plot FOO
I am going to filter the feature table for e.g. COTE adults from 2017 by hand, but then I just want to be able to use functions to calculate FOO and plot, to save on the amount of code I have to write.

```{r}

calc_FOO <- function(x, n_samples) {
  x %>% 
  mutate_if(is.numeric, ~1 * (. > 0)) %>%     # change to detection/non-detection
  summarise_each(funs = sum) %>%              # count number of detections
  melt() %>%                                  # make into long dataframe
  rename(Occurrence = value,
         Species = variable) %>% 
  mutate(FOO = Occurrence/n_samples*100)
}


# get rid of periods in column names and order species
scrub_periods <- function(x) {
  x$Species <-  gsub("\\.", " ", x$Species)
  x$Species <- factor(x$Species, levels = Species_ordered)
}

cols <- c(rgb(142, 202, 230, max = 255), "#8ECAE6", "#219EBC", "#023047", "#FFB703", "#FB8500")

# function to plot, with the species in the order determined by Species_with_colours
plot_FOO <- function(x) {
  x %>% 
    ggplot() +
    geom_bar(aes(x = Species, y = FOO), stat = "identity", fill = cols[1]) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45,  hjust=1)) +
    labs(y = "Frequency of occurrence (%)") +
    ylim(c(0, 100))
    #scale_fill_manual(values = cols[1])
}

```

### Monomoy chicks
Select the Monomoy chick data using ```filter```, change to presence/absence data using the ```mutate_if``` term, then ```melt``` to calculate the frequence of occurrence.

```{r, message=FALSE}
# Filter the data and save to object
MON_chicks <- df %>% 
  filter(Colony == "Monomoy NWR" & Age == "CHICK") %>% 
  select(Haddock:Carangidae.indet) 

# record number of records for calculating FOO
n_samples <- nrow(MON_chicks)
FOO <- calc_FOO(MON_chicks, n_samples)

FOO$Species <- scrub_periods(FOO)

FOO %>% 
  filter(FOO > 0) %>% 
  plot_FOO()
```


### Plotting them all together on one, multipanel figure

I think I can use faceting to do this, but I need to combine all the FOO data together into one dataframe. First, save the FOO data for each group.


```{r}

do_everything <- function(Colony_filter, Age_filter){
  x <- df %>% 
    filter(Colony == Colony_filter & Age == Age_filter) %>% 
    select(Haddock:Carangidae.indet)
  n_samples <- nrow(x)
  y <- calc_FOO(x, n_samples)
  y$Species <- scrub_periods(y)
  return(y)
}

FOO_MON_CHICK <- do_everything("Monomoy NWR", "CHICK") %>% mutate(Colony = "Monomoy NWR", Age = "CHICK")
FOO_MON_ADULT <- do_everything("Monomoy NWR", "ADULT") %>% mutate(Colony = "Monomoy NWR", Age = "ADULT")
FOO_SEA_CHICK <- do_everything("White and Seavey Is.", "CHICK") %>% mutate(Colony = "White and Seavey Is.", Age = "CHICK")
FOO_SEA_ADULT <- do_everything("White and Seavey Is.", "ADULT") %>% mutate(Colony = "White and Seavey Is.", Age = "ADULT")
FOO_STI_CHICK <- do_everything("Stratton Is.", "CHICK") %>% mutate(Colony = "Stratton Is.", Age = "CHICK")
FOO_STI_ADULT <- do_everything("Stratton Is.", "ADULT") %>% mutate(Colony = "Stratton Is.", Age = "ADULT")
FOO_OGI_CHICK <- do_everything("Outer Green Is.", "CHICK") %>% mutate(Colony = "Outer Green Is.", Age = "CHICK")
FOO_OGI_ADULT <- do_everything("Outer Green Is.", "ADULT") %>% mutate(Colony = "Outer Green Is.", Age = "ADULT")
FOO_JEN_CHICK <- do_everything("Jenny Is.", "CHICK") %>% mutate(Colony = "Jenny Is.", Age = "CHICK")
FOO_JEN_ADULT <- do_everything("Jenny Is.", "ADULT") %>% mutate(Colony = "Jenny Is.", Age = "ADULT")
FOO_PON_CHICK <- do_everything("Pond Is.", "CHICK") %>% mutate(Colony = "Pond Is.", Age = "CHICK")
FOO_PON_ADULT <- do_everything("Pond Is.", "ADULT") %>% mutate(Colony = "Pond Is.", Age = "ADULT")
FOO_EGG_CHICK <- do_everything("Eastern Egg Rock", "CHICK") %>% mutate(Colony = "Eastern Egg Rock", Age = "CHICK")
FOO_EGG_ADULT <- do_everything("Eastern Egg Rock", "ADULT") %>% mutate(Colony = "Eastern Egg Rock", Age = "ADULT")
FOO_MET_CHICK <- do_everything("Metinic", "CHICK") %>% mutate(Colony = "Metinic", Age = "CHICK")
FOO_MET_ADULT <- do_everything("Metinic", "ADULT") %>% mutate(Colony = "Metinic", Age = "ADULT")
FOO_SINWR_CHICK <- do_everything("Seal Is. NWR", "CHICK") %>% mutate(Colony = "Seal Is. NWR", Age = "CHICK")
FOO_SINWR_ADULT <- do_everything("Seal Is. NWR", "ADULT") %>% mutate(Colony = "Seal Is. NWR", Age = "ADULT")
FOO_SHI_CHICK <- do_everything("Ship Is.", "CHICK") %>% mutate(Colony = "Ship Is.", Age = "CHICK")
FOO_SHI_ADULT <- do_everything("Ship Is.", "ADULT") %>% mutate(Colony = "Ship Is.", Age = "ADULT")
FOO_PMI_CHICK <- do_everything("Petit Manan Is.", "CHICK") %>% mutate(Colony = "Petit Manan Is.", Age = "CHICK")
FOO_PMI_ADULT <- do_everything("Petit Manan Is.", "ADULT") %>% mutate(Colony = "Petit Manan Is.", Age = "ADULT")

All_FOO <- FOO_MON_CHICK %>% 
  bind_rows(., FOO_MON_ADULT) %>% 
  bind_rows(., FOO_SEA_CHICK) %>% 
  bind_rows(., FOO_SEA_ADULT) %>% 
  bind_rows(., FOO_STI_CHICK) %>% 
  bind_rows(., FOO_STI_ADULT) %>% 
  bind_rows(., FOO_OGI_CHICK) %>% 
  bind_rows(., FOO_OGI_ADULT) %>% 
  bind_rows(., FOO_JEN_CHICK) %>% 
  bind_rows(., FOO_JEN_ADULT) %>% 
  bind_rows(., FOO_PON_CHICK) %>%
  bind_rows(., FOO_PON_ADULT) %>%
  bind_rows(., FOO_EGG_CHICK) %>%
  bind_rows(., FOO_EGG_ADULT ) %>%
  bind_rows(., FOO_MET_CHICK) %>%
  bind_rows(., FOO_MET_ADULT) %>%
  bind_rows(., FOO_SINWR_CHICK) %>%
  bind_rows(., FOO_SINWR_ADULT) %>%
  bind_rows(., FOO_SHI_CHICK) %>%
  bind_rows(., FOO_SHI_ADULT) %>%
  bind_rows(., FOO_PMI_CHICK) %>%
  bind_rows(., FOO_PMI_ADULT) 
  
All_FOO$Species <- scrub_periods(All_FOO)

All_FOO %>% 
  filter(FOO > 0) %>% 
  ggplot() +
  geom_bar(aes(x = Species, y = FOO), stat = "identity") +
  facet_grid(rows = vars(Age), cols = vars(Colony)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,  hjust=1, colour = "black", margin = margin(t = 0)),
        axis.text.y = element_text(colour = "black"),
        axis.text=element_text(size=7),
        axis.title=element_text(size=8, face = "bold"),
        axis.ticks.x = element_blank(),
        panel.border=element_rect(colour="black",size=1, fill = NA),
        panel.spacing = unit(2, "mm"),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", size=8, face = "bold"),
        axis.line = element_blank()) +
  labs(y = "Frequency of occurrence (%)") +
  theme(legend.position = "none")
```

Great... there is way too much to try to plot this on one figure. Can I use pie charts instead?

```{r}
p <- All_FOO %>% 
  ggplot() +
  geom_bar(aes(x= Age, y = FOO, fill = Species, height = 1), stat = "identity", position = "fill") +
  facet_grid(~Colony) +
  scale_fill_manual(values = Species_with_colours) 
p


ggsave(filename = "MiFish/WSC3_terns/stacked_barplots_all_colonies.jpg",
  plot = p,
  dpi = 600,
  device = "jpeg",
  width = 10,
  height = 4
)

```

```{r}
All_FOO %>% group_by(Species) %>% 
  summarise(mean = mean(FOO))  %>% 
  arrange(-mean)

```


